---
output:
  html_document:
    toc: yes
    toc_depth: 2
    code_folding: hide
  word_document:
    toc: yes
    toc_depth: '2'
always_allow_html: yes
toc-title: Contents
editor_options:
  markdown:
    wrap: 72
---

```{r connect, include=FALSE}
library(RODBC)
library(kableExtra)
library(dplyr)
library(stringr)
library(readr)
library(ggplot2)
library(ggpubr)
library(tidyr)
library(sailr)
library(rmarkdown)
library(DT)
library(UpSetR)
library(patchwork)
library(gtsummary)
library(lubridate)
library(tidyr)
library(Cairo)
library(fastmap)
library(assertthat)
library(tidyverse)
library(gt)
library(gtsummary)
library(magrittr)
library(data.table)
library(dplyr)
library(broom)
library(purrr)
library(formattable)
library(readxl)
library(tableone)
library(openxlsx)
library(ComplexUpset)
options(scipen = 999)

```

```{r studywindow, echo=FALSE, message=FALSE}
study_start <- ymd("2000-01-01")
study_end   <- ymd("2022-12-31")

```

```{r log in, include = FALSE }
#log in to sail
source("")
#find_directory <- as.character(file.choose(""))
#source(find_directory)

#login to SAIL
login <- getLogin() #Username is optional - omit and it will prompt for username.
conn <- odbcConnect('PR_SAIL',login[1],login[2])
rm(login)

```


## SAIL Reviewer

Dear SAIL Reviewer,

Thank you for your time in reviewing this document.

We used individual-level data in this report for creation of a BMI
measure for the enitre population of Wales from 4 data sources along
with WDSD, ADDE and Ethnicity tables for general demographics data.

As indicated below this is version 7 of the file, where I have re-generated data from previous versions
for both adults and children and young people's cohort from 2000-2022 in accordance to feedback from journal reviewers.


To comply with disclosure rules we have removed rows or categories with
count < 10.


## Preamble

#### IGRP project 1151 (census data elements included)

#### Version 7

#### Study window

-   Start date: 2000-01-01
-   End date : 2021-12-31

#### Generated at/by

Compiled date: `r format(Sys.time(), '%a %d %B %Y')` Please note all
data coverage provided are up to above compiled date.

-   Michael Jeanne Childs [m.j.childs\@swansea.ac.uk](mailto:m.j.childs@swansea.ac.uk){.email}

#### Collaborators team

-   Ashley Akbari, Fatemeh Torabi, Sarah Aldridge, Hoda
    Abbasizanjani, Victoria Best, Helen Daniels, Gareth Davies, Ronan Lyons



## Results
Table X. Yearly characteristics for CYP (TableOne yearly)
```{r table1 for CYP yearly, eval=TRUE, echo=FALSE, message=FALSE}
table1_cyp <- paste0("
SELECT * FROM SAILW1151V.HDR25_NEW_BMI_TABLEONE_CYP
")

table1_run_cyp <- sqlQuery(conn, table1_cyp)

# changing labels and reordering factors 
table1_run_cyp$SEX[table1_run_cyp$SEX == 1] <- 'Male'
table1_run_cyp$SEX[table1_run_cyp$SEX == 2] <- 'Female'
table1_run_cyp <- table1_run_cyp %>%
              replace_na(list(ETHNICITY = 'Unknown'))

## creating the gtsummary
theme_gtsummary_compact()
table1_cyp_table1 <- 
  table1_run_cyp %>%
  select(BMI_YEAR, SEX, ETHNICITY, AGE_BAND, BMI_CAT, WIMD2019_QUINTILE, RURAL_URBAN) %>%
  filter(BMI_YEAR < 2011) %>%
    mutate(AGE_BAND = factor(AGE_BAND, levels = c('2-5', '5-13', '13-19')),
           BMI_CAT = factor(BMI_CAT, levels = c('Underweight', 'Normal weight', 'Obese', 'Overweight', 'Unknown'))) %>%
  tbl_summary(by = BMI_YEAR,
             statistic = list(all_continuous() ~ "{mean}, ({sd})"),
             label = list(SEX ~"Sex",
                          ETHNICITY ~ "Ethnicity",
                          AGE_BAND ~ "Age band",
                          BMI_CAT ~ "BMI category",
                          WIMD2019_QUINTILE ~ "Welsh Index of Multiple Deprivation 2019*",
                          RURAL_URBAN ~ "Rural urban classification")
                        ) %>%
  # add n 
  #add_overall() %>%
  bold_labels() %>%
  #tab_header("Characteristics of CYP population") %>%
   as_flex_table() %>%
    flextable::save_as_docx(path = 'P:/childsm/BMI - rechecking for inconsistent entries/TableOne_cyp.docx')

table1_cyp_table2 <- 
  table1_run_cyp %>%
  select(BMI_YEAR, SEX, ETHNICITY, AGE_BAND, BMI_CAT, WIMD2019_QUINTILE, RURAL_URBAN) %>%
  filter(BMI_YEAR > 2010) %>%
    mutate(AGE_BAND = factor(AGE_BAND, levels = c('2-5', '5-13', '13-19')),
           BMI_CAT = factor(BMI_CAT, levels = c('Underweight', 'Normal weight', 'Obese', 'Overweight', 'Unknown'))) %>%
  tbl_summary(by = BMI_YEAR,
             statistic = list(all_continuous() ~ "{mean}, ({sd})"),
             label = list(SEX ~"Sex",
                          ETHNICITY ~ "Ethnicity",
                          AGE_BAND ~ "Age band",
                          BMI_CAT ~ "BMI category",
                          WIMD2019_QUINTILE ~ "Welsh Index of Multiple Deprivation 2019*",
                          RURAL_URBAN ~ "Rural urban classification")
                        ) %>%
  # add n 
  #add_overall() %>%
  bold_labels() %>%
  #tab_header("Characteristics of CYP population") %>%
   as_flex_table() %>%
    flextable::save_as_docx(path = 'P:/childsm/BMI - rechecking for inconsistent entries/TableOne_cyp_cont.docx')
 
```

Table X. Yearly characteristics for Adults (TableOne yearly)
```{r table1 for adults yearly, eval=TRUE, echo=FALSE, message=FALSE}
table1_adult <- paste0("
SELECT * FROM SAILW1151V.HDR25_NEW_BMI_TABLEONE_ADULTS
")

table1_run_adult <- sqlQuery(conn, table1_adult)

# changing labels and reordering factors 
table1_run_adult$SEX[table1_run_adult$SEX == 1] <- 'Male'
table1_run_adult$SEX[table1_run_adult$SEX == 2] <- 'Female'
table1_run_adult <- table1_run_adult %>%
              replace_na(list(ETHNICITY = 'Unknown'))

## creating the gtsummary
theme_gtsummary_compact()
table1_adult_table1 <- 
  table1_run_adult %>%
  select(BMI_YEAR, SEX, ETHNICITY, AGE_BAND, BMI_CAT, WIMD2019_QUINTILE, RURAL_URBAN) %>%
  filter(BMI_YEAR < 2011) %>%
    mutate(BMI_CAT = factor(BMI_CAT, levels = c('Underweight', 'Normal weight', 'Obese', 'Overweight', 'Unknown'))) %>%
  tbl_summary(by = BMI_YEAR,
             statistic = list(all_continuous() ~ "{mean}, ({sd})"),
             label = list(SEX ~"Sex",
                          ETHNICITY ~ "Ethnicity",
                          AGE_BAND ~ "Age band",
                          BMI_CAT ~ "BMI category",
                          WIMD2019_QUINTILE ~ "Welsh Index of Multiple Deprivation 2019*",
                          RURAL_URBAN ~ "Rural urban classification")
                        ) %>%
  # add n 
  #add_overall() %>%
  bold_labels() %>%
  #tab_header("Characteristics of adult population") %>%
   as_flex_table() %>%
    flextable::save_as_docx(path = 'P:/childsm/BMI - rechecking for inconsistent entries/TableOne_adult.docx')

table1_adult_table2 <- 
  table1_run_adult %>%
  select(BMI_YEAR, SEX, ETHNICITY, AGE_BAND, BMI_CAT, WIMD2019_QUINTILE, RURAL_URBAN) %>%
  filter(BMI_YEAR > 2010) %>%
    mutate(BMI_CAT = factor(BMI_CAT, levels = c('Underweight', 'Normal weight', 'Obese', 'Overweight', 'Unknown'))) %>%
  tbl_summary(by = BMI_YEAR,
             statistic = list(all_continuous() ~ "{mean}, ({sd})"),
             label = list(SEX ~"Sex",
                          ETHNICITY ~ "Ethnicity",
                          AGE_BAND ~ "Age band",
                          BMI_CAT ~ "BMI category",
                          WIMD2019_QUINTILE ~ "Welsh Index of Multiple Deprivation 2019*",
                          RURAL_URBAN ~ "Rural urban classification")
                        ) %>%
  # add n 
  #add_overall() %>%
  bold_labels() %>%
  #tab_header("Characteristics of adult population") %>%
   as_flex_table() %>%
    flextable::save_as_docx(path = 'P:/childsm/BMI - rechecking for inconsistent entries/TableOne_adult_cont.docx')
 
table1_adult_table2
```

Table X. Yearly characteristics for CYP (TableOne 5 yearly)
```{r table2 for CYP 5 yearly, eval=TRUE, echo=FALSE, message=FALSE}
table2_cyp <- paste0("
SELECT ALF_E, BMI_YEAR, SEX, ETHNICITY, AGE_BAND, BMI_CAT, WIMD2019_QUINTILE, RURAL_URBAN FROM 
SAILW1151V.HDR25_NEW_BMI_CLEAN_CYP_POP_5YEARLY
")

table2_run_cyp <- sqlQuery(conn, table2_cyp)

# changing labels and reordering factors 
table2_run_cyp$SEX[table2_run_cyp$SEX == 1] <- 'Male'
table2_run_cyp$SEX[table2_run_cyp$SEX == 2] <- 'Female'
table2_run_cyp <- table2_run_cyp %>%
              replace_na(list(ETHNICITY = 'Unknown'))

## creating the gtsummary
theme_gtsummary_compact()
table2_cyp_table <- table2_run_cyp %>%
    mutate(AGE_BAND = factor(AGE_BAND, levels = c('2-5', '5-13', '13-19')),
           BMI_CAT = factor(BMI_CAT, levels = c('Underweight', 'Normal weight', 'Obese', 'Overweight', 'Unknown'))) %>%
  tbl_summary(include =c(BMI_YEAR, SEX, ETHNICITY, AGE_BAND, BMI_CAT, WIMD2019_QUINTILE, RURAL_URBAN),
              by = BMI_YEAR,
             statistic = list(all_continuous() ~ "{mean}, ({sd})"),
             label = list(SEX ~"Sex",
                          ETHNICITY ~ "Ethnicity",
                          AGE_BAND ~ "Age band",
                          BMI_CAT ~ "BMI category",
                          WIMD2019_QUINTILE ~ "Welsh Index of Multiple Deprivation 2019*",
                          RURAL_URBAN ~ "Rural urban classification")
                        ) %>%
  bold_labels() %>%
  #tab_header("Characteristics of CYP population") %>%
   as_flex_table() %>%
    flextable::save_as_docx(path = 'P:/childsm/BMI - rechecking for inconsistent entries/TableOne_5years_cyp.docx')
 
```

Table X. Yearly characteristics for Adults (TableOne 5 yearly)
```{r table2 for adults 5 yearly, eval=TRUE, echo=FALSE, message=FALSE}
table2_adults <- paste0("
SELECT ALF_E, BMI_YEAR, SEX, ETHNICITY, AGE_BAND, BMI_CAT, WIMD2019_QUINTILE, RURAL_URBAN FROM SAILW1151V.HDR25_NEW_BMI_CLEAN_adults_POP_5YEARLY
")

table2_run_adults <- sqlQuery(conn, table2_adults)

# changing labels and reordering factors 
table2_run_adults$SEX[table2_run_adults$SEX == 1] <- 'Male'
table2_run_adults$SEX[table2_run_adults$SEX == 2] <- 'Female'
table2_run_adults <- table2_run_adults %>%
              replace_na(list(ETHNICITY = 'Unknown'))

## creating the gtsummary
theme_gtsummary_compact()
table2_adults_table <-  table2_run_adults %>%
    mutate(BMI_CAT = factor(BMI_CAT, levels = c('Underweight', 'Normal weight', 'Obese', 'Overweight', 'Unknown'))) %>%
  tbl_summary(include =c(BMI_YEAR, SEX, ETHNICITY, AGE_BAND, BMI_CAT, WIMD2019_QUINTILE, RURAL_URBAN),
              by = BMI_YEAR,
             statistic = list(all_continuous() ~ "{mean}, ({sd})"),
             label = list(SEX ~"Sex",
                          ETHNICITY ~ "Ethnicity",
                          AGE_BAND ~ "Age band",
                          BMI_CAT ~ "BMI category",
                          WIMD2019_QUINTILE ~ "Welsh Index of Multiple Deprivation 2019*",
                          RURAL_URBAN ~ "Rural urban classification")
                        ) %>%
  # add n 
  #add_overall() %>%
  bold_labels() %>%
  #tab_header("Characteristics of adults population") %>%
   as_flex_table() %>%
    flextable::save_as_docx(path = 'P:/childsm/BMI - rechecking for inconsistent entries/TableOne_5years_adults.docx')
 
```

**Figure X.** Trends of known BMI - CYP and Adults --- REGENERATED AND SAVED.
```{r trend_overall, echo=FALSE, message=FALSE, fig.width=12, fig.height=10, dpi = 300}
# saving all BMI counts into.csv
all_bmi_counts <- paste0("SELECT * FROM SAILW1151V.HDR25_NEW_BMI_LINEPLOT_ALL")
all_bmi_counts_run <- sqlQuery (conn, all_bmi_counts)
# save to .csv
write.csv(all_bmi_counts_run, file = 'all_bmi_counts.csv')

##################################
## creating the line plot
##################################

bmi_all_trend_plot <-  ggplot(all_bmi_counts_run %>%
                               filter(BMI_CAT %in% c("Underweight", "Normal weight", "Overweight", "Obese")) %>%
                               mutate(POP = factor(POP, levels = c("CYP", "Adults"))), 
         aes(x=BMI_YEAR, y=FREQUENCY, group=factor(BMI_CAT, levels = c("Underweight", "Normal weight", "Overweight", "Obese"))))+
  geom_line(alpha = 0.5, linewidth = 1) +
  geom_point(size = 4, aes(shape=factor(BMI_CAT, levels = c("Underweight", "Normal weight", "Overweight", "Obese")))) +
  facet_wrap(~POP, ncol = 1, scales = "free_y")+
  scale_x_continuous(breaks = 2000:2022) +  
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x)))+
  #ylim(0, 30000) +
  labs(x = 'Year', y = 'Frequency') +
  scale_shape_discrete(name = 'BMI Categories') +
  theme_light() +
  theme(legend.position = "bottom",) +
  theme(axis.text.x = element_text(size=12, angle=45, hjust= .5, vjust = .5),
        axis.text.y = element_text(size=12),
        axis.title = element_text(size= 15),
        axis.title.y = element_text(size= 15),
        axis.title.x = element_text(size= 15),
        strip.text = element_text(size= 15))

# 4. save as image
ggsave(filename = 'trend_overall.png', width = 10, height = 7, device = "tiff", dpi = 500)


bmi_all_trend_plot
```

**Figure 3.** Data source distribution for population with BMI readings.
```{r upsetplot_all, include = FALSE}
## this chunk creates an upset plot for all, cyp, and adult cohorts. We then stitch them onto one image and that's what we will load in the end.

##loading upset plot for adults
upsetplot_adults <- paste0("SELECT * FROM SAILW1151V.HDR25_NEW_BMI_UPSETPLOT_ADULTS")
upsetplot_adults_run <- sqlQuery(conn, upsetplot_adults)

png(file="upset_adults.png", width=1000, height = 400, units="px")

# renaming columns to how we want them to look like in the plot:
colnames(upsetplot_adults_run) <- c('ALF_E', 'WLGP', 'MIDS', 'PEDW')

# identifying the sets/columns to use
flags_upset_adults = colnames(upsetplot_adults_run)
# this changes the column into TRUE/FALSE flags for the package to run.
upsetplot_adults_run[flags_upset_adults] = upsetplot_adults_run[flags_upset_adults] == 1
# this allows you to see what this subset looks like:
t(head(upsetplot_adults_run, 3))
```


```{r upsetplot_all, include = FALSE}
## this chunk creates an upset plot for all, cyp, and adult cohorts. We then stitch them onto one image and that's what we will load in the end.

##loading upset plot for all
##loading upset plot for adults
upsetplot_cyp <- paste0("SELECT * FROM SAILW1151V.HDR25_NEW_BMI_UPSETPLOT_cyp")
upsetplot_cyp_run <- sqlQuery(conn, upsetplot_cyp)

# renaming columns to how we want them to look like in the plot:
colnames(upsetplot_cyp_run) <- c('ALF_E', 'WLGP', 'MIDS', 'PEDW', 'NCCH')

# identifying the sets/columns to use
flags_upset_cyp = colnames(upsetplot_cyp_run)
# this changes the column into TRUE/FALSE flags for the package to run.
upsetplot_cyp_run[flags_upset_cyp] = upsetplot_cyp_run[flags_upset_cyp] == 1
# this allows you to see what this subset looks like:
t(head(upsetplot_cyp_run, 3))
```

```{r upset_plots_all_checks, echo = FALSE, message = FALSE, warning = TRUE, quiet = TRUE,  figure.height = 5, figure.width = 8}
## this was run as part of the supplementary material requested by SAIL reviewer to ensure that each group follows the HESA guidelines.
## creating upset plot for the overall crcs
upset_adult_bmi <- upset(upsetplot_adults_run, flags_upset_adults, 
      # labels the x-axis
      name='Data source', 
      # changes the size of the upset plot:
      width_ratio = 0.1,
      height_ratio= 0.5,
      stripes='white',
      min_size=100,
      themes=upset_modify_themes(
        list('intersections_matrix'=theme(axis.text.y=element_text(color='black')))),
      base_annotations=list(
        
        'Intersection size (%)' = 
          intersection_size(mapping = aes(fill='bars_color'),
                            text=list(angle=0, size=3)
                            #this changes the label on the intersection size to %
                           #text_mapping=aes(label=paste0(round(!!get_size_mode('exclusive_intersection')/nrow(upset_plot_all_children_run) * 100), '%')))
          )
        + scale_y_continuous(
              labels=scales::percent_format(scale=100/nrow(upsetplot_adults_run)),
              breaks=c(0,20, 40, 60, 80)/ 100 * nrow(upsetplot_adults_run))
        + scale_fill_manual(values=c('bars_color'='black'), guide='none')
        + theme(axis.text.y=element_text(color = 'black'))
      ),
     queries=list(
      # this sets the color for each type of disability - this matches the colors from the horizontal bar plot.
       upset_query(set = 'WLGP' , color = '#34eb3a', fill = '#34eb3a'),
       upset_query(set = 'MIDS' , color = '#E69F00', fill = '#E69F00'),
       upset_query(set = 'PEDW' , color = '#56B4E9', fill = '#56B4E9')
     ),
     set_sizes=(
       upset_set_size(
         #this changes the label on the intersection size to %
         #mapping=aes(y=..count../nrow(upsetplot_adults_run) * 100, '%')) -- older version
         mapping=aes(y=after_stat(count)/nrow(upsetplot_adults_run) * 100, '%'))
       )
       + ylab('Category \nsize (%)')
       + theme(axis.text.x=element_text(angle=90, hjust=.1, vjust=.1, color = 'black'))
       + theme(axis.ticks.x = element_line())
     + ggtitle ('Adults')
     )

# saving all children with disabilities in CRCS plot
ggsave('upset_adult_bmi.jpeg', upset_adult_bmi, device = 'jpeg', height = 150, width = 250, unit = 'mm', dpi = 400) 

upset_adult_bmi

##png::writePNG(img123,"P:/childsm/BMI/bmi_alg_report/upset_all_three.png" )
```

```{r upset_plots_all_checks, echo = FALSE, message = FALSE, warning = TRUE, quiet = TRUE,  figure.height = 5, figure.width = 8}
## this was run as part of the supplementary material requested by SAIL reviewer to ensure that each group follows the HESA guidelines.
## creating upset plot for the overall crcs
upset_cyp_bmi <- upset(upsetplot_cyp_run, flags_upset_cyp, 
      # labels the x-axis
      name='Data source', 
      # changes the size of the upset plot:
      width_ratio = 0.1,
      height_ratio= 0.5,
      stripes='white',
      min_size=100,
      themes=upset_modify_themes(
        list('intersections_matrix'=theme(axis.text.y=element_text(color='black')))),
      base_annotations=list(
        
        'Intersection size (%)' = 
          intersection_size(mapping = aes(fill='bars_color'),
                            text=list(angle=0, size=3)
                            #this changes the label on the intersection size to %
                           #text_mapping=aes(label=paste0(round(!!get_size_mode('exclusive_intersection')/nrow(upset_plot_all_children_run) * 100), '%')))
          )
        + scale_y_continuous(
              labels=scales::percent_format(scale=100/nrow(upsetplot_cyp_run)),
              breaks=c(0,20, 40, 60, 80)/ 100 * nrow(upsetplot_cyp_run))
        + scale_fill_manual(values=c('bars_color'='black'), guide='none')
        + theme(axis.text.y=element_text(color = 'black'))
      ),
     queries=list(
      # this sets the color for each type of disability - this matches the colors from the horizontal bar plot.
       upset_query(set = 'WLGP' , color = '#34eb3a', fill = '#34eb3a'),
       upset_query(set = 'MIDS' , color = '#E69F00', fill = '#E69F00'),
       upset_query(set = 'PEDW' , color = '#56B4E9', fill = '#56B4E9'),
       upset_query(set = 'NCCH' , color = 'purple', fill = 'purple')
     ),
     set_sizes=(
       upset_set_size(
         #this changes the label on the intersection size to %
         mapping=aes(y=after_stat(count)/nrow(upsetplot_cyp_run) * 100, '%'))
       )
       + ylab('Category \nsize (%)')
       + theme(axis.text.x=element_text(angle=90, hjust=.1, vjust=.1, color = 'black'))
       + theme(axis.ticks.x = element_line())
     + ggtitle ('CYP')
     )

# 


# use patchwork to put together the two upset plotsggsave('upset_cyp_bmi.jpeg', upset_cyp_bmi, device = 'jpeg', height = 150, width = 250, unit = 'mm', dpi = 400) 

upset_all <- upset_cyp_bmi / upset_adult_bmi

upset_all

ggsave('upset_all_bmi.jpeg', upset_all, device = 'jpeg', height = 200, width = 250, unit = 'mm', dpi = 400) 


```

**Figure X.** Preserving BMI records for cyp and adult population.
```{r stacked plots adults, echo = FALSE, message = FALSE, fig.width = 12, fig.height = 8, dpi = 300}
## using facet to create stacked plots
all_three <- paste0("
select * from SAILW1151V.HDR25_NEW_BMI_PRESERVATION_TABLE
")

all_three_run <- sqlQuery(conn, all_three)

# changing the names of lookback values
all_three_run$LOOKBACK <- str_replace_all(all_three_run$LOOKBACK, 'all years', 'All years')

# calculating the percentage for each lookback period, bmi_year, bmi_month, and bmi_cat
counts_all_three <- all_three_run %>%
  group_by(POP, LOOKBACK, BMI_YEAR, BMI_MONTH, BMI_CAT) %>%
  summarise(n = sum(COUNTS)) %>%
  mutate(perc = n / sum(n) * 100)

# save to csv.
write.csv(counts_all_three, file = 'preservation_table.csv')

preserved_all <- ggplot(counts_all_three %>%
                          # specifying the order of variables to appear on the graph
                            mutate(BMI_MONTH = factor(BMI_MONTH, levels = c("Jan-2005", "Feb-2005", "Mar-2005", "Apr-2005", "May-2005", "Jun-2005", "Jul-2005", "Aug-2005", "Sep-2005", "Oct-2005", "Nov-2005", "Dec-2005",
                                                                            "Jan-2006", "Feb-2006", "Mar-2006", "Apr-2006", "May-2006", "Jun-2006", "Jul-2006", "Aug-2006", "Sep-2006", "Oct-2006", "Nov-2006", "Dec-2006",
                                                                            "Jan-2007", "Feb-2007", "Mar-2007", "Apr-2007", "May-2007", "Jun-2007", "Jul-2007", "Aug-2007", "Sep-2007", "Oct-2007", "Nov-2007", "Dec-2007",
                                                                            "Jan-2008", "Feb-2008", "Mar-2008", "Apr-2008", "May-2008", "Jun-2008", "Jul-2008", "Aug-2008", "Sep-2008", "Oct-2008", "Nov-2008", "Dec-2008",
                                                                            "Jan-2009", "Feb-2009", "Mar-2009", "Apr-2009", "May-2009", "Jun-2009", "Jul-2009", "Aug-2009", "Sep-2009", "Oct-2009", "Nov-2009", "Dec-2009",
                                                                            "Jan-2010", "Feb-2010", "Mar-2010", "Apr-2010", "May-2010", "Jun-2010", "Jul-2010", "Aug-2010", "Sep-2010", "Oct-2010", "Nov-2010", "Dec-2010",
                                                                            "Jan-2011", "Feb-2011", "Mar-2011", "Apr-2011", "May-2011", "Jun-2011", "Jul-2011", "Aug-2011", "Sep-2011", "Oct-2011", "Nov-2011", "Dec-2011",
                                                                            "Jan-2012", "Feb-2012", "Mar-2012", "Apr-2012", "May-2012", "Jun-2012", "Jul-2012", "Aug-2012", "Sep-2012", "Oct-2012", "Nov-2012", "Dec-2012",
                                                                            "Jan-2013", "Feb-2013", "Mar-2013", "Apr-2013", "May-2013", "Jun-2013", "Jul-2013", "Aug-2013", "Sep-2013", "Oct-2013", "Nov-2013", "Dec-2013",
                                                                            "Jan-2014", "Feb-2014", "Mar-2014", "Apr-2014", "May-2014", "Jun-2014", "Jul-2014", "Aug-2014", "Sep-2014", "Oct-2014", "Nov-2014", "Dec-2014",
                                                                            "Jan-2015", "Feb-2015", "Mar-2015", "Apr-2015", "May-2015", "Jun-2015", "Jul-2015", "Aug-2015", "Sep-2015", "Oct-2015", "Nov-2015", "Dec-2015",
                                                                            "Jan-2016", "Feb-2016", "Mar-2016", "Apr-2016", "May-2016", "Jun-2016", "Jul-2016", "Aug-2016", "Sep-2016", "Oct-2016", "Nov-2016", "Dec-2016",
                                                                            "Jan-2017", "Feb-2017", "Mar-2017", "Apr-2017", "May-2017", "Jun-2017", "Jul-2017", "Aug-2017", "Sep-2017", "Oct-2017", "Nov-2017", "Dec-2017",
                                                                            "Jan-2018", "Feb-2018", "Mar-2018", "Apr-2018", "May-2018", "Jun-2018", "Jul-2018", "Aug-2018", "Sep-2018", "Oct-2018", "Nov-2018", "Dec-2018",
                                                                            "Jan-2019", "Feb-2019", "Mar-2019", "Apr-2019", "May-2019", "Jun-2019", "Jul-2019", "Aug-2019", "Sep-2019", "Oct-2019", "Nov-2019", "Dec-2019",
                                                                            "Jan-2020", "Feb-2020", "Mar-2020", "Apr-2020", "May-2020", "Jun-2020", "Jul-2020", "Aug-2020", "Sep-2020", "Oct-2020", "Nov-2020", "Dec-2020",
                                                                            "Jan-2021", "Feb-2021", "Mar-2021", "Apr-2021", "May-2021", "Jun-2021", "Jul-2021", "Aug-2021", "Sep-2021", "Oct-2021", "Nov-2021", "Dec-2021",
                                                                            "Jan-2022", "Feb-2022", "Mar-2022", "Apr-2022", "May-2022", "Jun-2022", "Jul-2022", "Aug-2022", "Sep-2022", "Oct-2022", "Nov-2022", "Dec-2022")),
                                   LOOKBACK = factor(LOOKBACK, levels = c("1 year", "5 years", "All years")),
                                   POP = factor(POP, levels = c("CYP", "Adults")),
                                   BMI_CAT = factor(BMI_CAT, levels = c('Underweight', 'Normal weight', 'Overweight', 'Obese', 'Unknown'))), 
                          aes(x = BMI_MONTH, y = perc, fill = BMI_CAT)) +
  geom_col(alpha = 0.6, linewidth =1) +
  facet_grid(POP ~LOOKBACK, scales = "free") +
  xlab("Month and Year") +
  ylab("Percentage (%)") +
  scale_x_discrete(name = 'Month and Year', breaks=c("Jan-2005", 
                                                     "Jan-2010",
                                                     "Jan-2015",
                                                     "Jan-2020")) +
  scale_fill_manual(name = "BMI Categories", breaks =c('Underweight', 'Normal weight', 'Overweight', 'Obese', 'Unknown'),
                    values =c('black', 'green', 'purple', 'red', 'gray')) +
  theme(legend.position = "bottom") +
  theme_light() +
  theme(axis.text.x = element_text(size=7, angle=45, hjust = .5, vjust = .5))


library(ggthemes)
#saving file
ggsave(filename = 'preseved_all.png', width = 10, height = 7, device = "tiff", dpi = 500 )

preserved_all


```





## Supplementary Material

**A.** Read codes pertaining to BMI related events in WLGP database
```{r bmi_code, echo = FALSE, message=FALSE}
# Code list
bmi_lookup <- data.frame (bmi_code = c("2293","229..","229Z.","2292","2294","2295","2291","22A..","22A1.","22A2.","22A3.","22A4.","22A5.","22A6.","22AA.","22AZ.","1266","1444","22K3.","22K..","22K1.","22K2.","22K4.","22K5.","22K6.","22K7.","22K8.","22K9.","22KC.","22KC.","22KD.","22KD.","22KE.","22KE.","66C4.","66C6.","66CE.","8CV7.","8T11.","C38..","C380.","C3800","C3801","C3802","C3803","C3804","C3805","C3806","C3807","C38z.","C38z0","Cyu7.","22K4.","22A1.","22A2.","22A3.","22A4.","22A5.","22A6.","22AA.","R0348","66C1.","66C2.","66C5.","66CX.","66CZ.","9hN..","9OK..","9OK1.","9OK3.","9OK2.","9OK4.","9OK5.","9OK6.","9OK7.","9OK8.","9OKA.","9OKZ.","C38y0"),
             description = c("O/E -height within 10% average","O/E - height","O/E - height NOS","O/E - height 10-20% < average","O/E-height 10-20% over average","O/E -height > 20% over average","O/E-height > 20% below average","O/E - weight","O/E - weight > 20% below ideal","O/E -weight 10-20% below ideal","O/E - weight within 10% ideal","O/E - weight 10-20% over ideal","O/E - weight > 20% over ideal","O/E - Underweight","Overweight","O/E - weight NOS","FH: Obesity","H/O: obesity","Body Mass Index low K/M2","Body Mass Index","Body Mass Index normal K/M2","Body Mass Index high K/M2","Body mass index index 25-29 - overweight","Body mass index 30+ - obesity","Body mass index less than 20","Body mass index 40+ - severely obese","Body mass index 20-24 - normal","Body mass index centile","Obese class I (body mass index 30.0 - 34.9)","Obese class I (BMI 30.0-34.9)","Obese class II (body mass index 35.0 - 39.9)","Obese class II (BMI 35.0-39.9)","Obese class III (BMI equal to or greater than 40.0)","Obese cls III (BMI eq/gr 40.0)","Has seen dietician - obesity","Treatment of obesity started","Reason for obesity therapy - occupational","Anti-obesity drug therapy commenced","Rfrrl multidisip obesity clin","Obesity/oth hyperalimentation","Obesity","Obesity due to excess calories","Drug-induced obesity","Extrem obesity+alveol hypovent","Morbid obesity","Central obesity","Generalised obesity","Adult-onset obesity","Lifelong obesity","Obesity/oth hyperalimentat NOS","Simple obesity NOS","[X]Obesity+oth hyperalimentatn","BMI 25-29 - overweight","O/E - weight > 20% below ideal","O/E -weight 10-20% below ideal","O/E - weight within 10% ideal","O/E - weight 10-20% over ideal","O/E - weight > 20% over ideal","O/E - Underweight","Overweight","[D] Underweight","Itinital obesity assessment","Follow-up obesity assessment","Treatment of obesity changed","Obesity multidisciplinary case review","Obesity monitoring NOS","Exception reporting: obesity quality indicators","Obesity monitoring admin.","Attends obesity monitoring","Obesity monitoring default","Refuses obesity monitoring","Obesity monitoring 1st letter","Obesity monitoring 2nd letter","Obesity monitoring 3rd letter","Obesity monitoring verbal inv.","Obesity monitor phone invite","Obesity monitoring check done","Obesity monitoring admin.NOS","Pickwickian syndrome"),
             complexity = c("where event_val between x and y (depending on unit)","where event_val between x and y (depending on unit)","where event_val between x and y (depending on unit)","height","height","height","height","where event_val between 32 and 250","where event_val between 32 and 250","where event_val between 32 and 250","where event_val between 32 and 250","where event_val between 32 and 250","where event_val between 32 and 250","where event_val between 32 and 250","where event_val between 32 and 250","where event_val between 32 and 250","Obese","Obese","Underweight","bmi","Normal weight","Overweight/Obese","Overweight","Obese","Underweight","Obese","Normal weight","bmi","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Overweight","Underweight","Underweight","Normal weight","Overweight","Overweight","Underweight","Overweight","Underweight","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese","Obese"),
             category = c("height","height","height","height","height","height","height","weight","weight","weight","weight","weight","weight","weight","weight","weight","obese","obese","underweight","bmi","normal weight","obese","overweight","obese","underweight","obese","normal weight","bmi","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","overweight","underweight","underweight","normal weight","overweight","overweight","underweight","overweight","underweight","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese","obese")
            )

DT::datatable(bmi_lookup, rownames = FALSE, extensions=c("Buttons"),options = list(dom= "Bfrtip", buttons = c("copy","csv")))
```
**B.** ICD-10 codes pertaining to BMI related events in PEDW database
```{r ICD-10 codes, echo = FALSE, message=FALSE}
icd_10 <- data.frame (icd_code = c("E66.0", "E66.1", "E66.2", "E66.8", "E66.9"),
          description = c("Obesity due to excess calories", "Drug-induced obesity (use additional external cause code (Chapter XX, if desired, to identify the drug)", "Extreme obesity with alveolar hypoventilation (Obesity hypoventilation syndrome / Pickwickian syndrome)", "Other obesity (Morbid obesity)", "Obesity, unspecified (Simple obesity, NOS)"),
          category = c("obese", "obese", "obese", "obese", "obese")
          )

DT::datatable(icd_10, rownames = FALSE, extensions=c("Buttons"),options = list(dom= "Bfrtip", buttons = c("copy","csv")))
```


## Appendix

Code created by: Michael Jeanne Childs, m.j.childs\@swansea.ac.uk
Based from code by: Sarah Aldridge

This folder contains rmd file for report of BMI algorithm
methodology (which contains script used to create dataframes for results) data needed
to create tables and figures for the report. Rows with counts \< 10 were not included.

SAIL tables used to generate the report and tables are as follows:

1\. SAILWMC_V.C19_COHORT_WLGP_GP_EVENT_CLEANSED - to extract ALFs with
BMI related events

2\. SAILWMC_V.C19_COHORT_PEDW_SPELL - to extract ALFs with Obesity
related hospital admissions

3\. SAILWMC_V.C19_COHORT_PEDW_DIAG - to extract ALFs with Obesity
related hospital admissions

4\. SAILWMC_V.C19_COHORT_MIDS_INITIAL_ASSESSMENT - to extract ALFs with
height and weight values

5\. SAILWMC_V.C19_COHORT_MIDS_BIRTH - to link birth dates to
MOTHER_ALF_PEs and calculate 9months before/after birthdate to allocate
post/pre-natal related weight changes.

6\. SAILWMC_V.C19_COHORT_NCCH_CHILD_MEASUREMENT_PROGRAM - to extract
ALFs with BMI related events

7\. SAILWMC_V.C19_COHORT_NCCH_EXAM- to extract ALFs with BMI related
events

8\. SAILWMC_V.C19_COHORT_NCCH_CHILD_BIRTHS - used to link NCCH table to
tables from other databases.

9\. sailwmc_v.C19_COHORT_WDSD_PER_RESIDENCE_GPREG - to extract week of
birth, sex, date of death, and residency information from ALFs to
calculate age at BMI recording and assign age bands to ALFs

10\. SAILW1151V.RRDA_ETHN - to extract Ethnicity information

11\. SAILWMC_V.C19_COHORT_WDSD_CLEAN_ADD_GEOG_CHAR_LSOA2011 - to extract
WIMD2019_QUINTILE information for ALFs

12\. sailw1151v.HDR25_BMI_GEOGRAPHY_LOOKUP - to extract rural/urban
classification and local health board information

13\. sailwmc_v.C19_COHORT_ADDE_DEATHS -- to extract date of death for
    individuals.

| **Document name** | **Description**                                                                                              |
|:----------------------------------|:------------------------------------|
| BMI_R_Report.Rmd  | R file with details on how the BMI algorithm runs and has reports on the results of the algorithm. |
| BMI_R_Report.html   | R markdown rendered to html for easier reading.                                                        |

### Data subfolder

| **Document name**     | **Description**   |                                                                         
|:----------------------|:------------------|
| Figures.docx          | Word document version with all the figures created from the code. |
| TableOnes.docx        | Word document version of the gtsummary table created from the R code. Has the CYP and adult versions for easier transfer when writing a manuscript              |

### Figures subfolder

| **Document name**        | **Description**                                                                                                                                                                                  |
|:----------------------------------|:------------------------------------|
| flowchart.png        | Diagram showing the process used in the methodology with counts for each step.                                                                                                                                             |
| flowchart_methods.jpg   | Diagram showing the methodology |
| preserved_adults.png      | Shows stacked plots of the proportion of BMI categories retrieved when we preserve BMI readings from 1 year or 5 years before study end date, or the whole of the study period for adult cohort. |
| preserved_CYP.png         | Shows stacked plots of the proportion of BMI categories retrieved when we preserve BMI readings from 1 year or 5 years before study end date, or the whole of the study period for CYP cohort.   |
| stacked_adults.png    | Line graph showing the trend of counts of known and unknown BMI categories from 2010-2022 in the adult cohort grouped by data source.                                                            |
| stacked_cyp.png       | Line graph showing the trend of counts of known and unknown BMI categories from 2010-2022 in the CYP cohort grouped by data source.                                                              |
| trend_known.png         | Line graph showing the trend of counts of known BMI categories from 2000-2022 for the CYP and adult cohort. x-axis = year, y-axis = frequency                                                    |
| trend_overall.png       | Line graph showing the trend of counts of known BMI categories from 2000-2022 for the CYP and adult cohort. x-axis = year, y-axis = frequency                                                    |
| trend_with_unknown.png  | Line graph showing the trend of counts of known and unknown BMI categories from 2000-2022 for the CYP and adult cohort. x-axis = year, y-axis = frequency                                        |
| upset_adults.png | Plot showing the number of individuals in adult population cohort with BMI records from a single data source and overlaps, and those with entries from either two, or three data sources.        |
| upset_all.png    | Plot showing the number of individuals in whole population cohort with BMI records from a single data source and overlaps, and those with entries from either two, three, or four data sources.  |
| upset_all_three.png    | All three upset plots patched in one file |
| upset_cyp.png    | Plot showing the number of individuals in CYP population cohort with BMI records from a single data source and overlaps, and those with entries from either two, three, or four data sources.    |
